NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2017-12-20 Manuscript/manuscriptPS.tex
Update figures and some text

\subsection{PS headgroup and glycerol backbone structure in simulations and experiments}
The headgroup order parameters of DOPS and POPS bilayers 
from different experiments and simulations are shown in
Figs. \ref{HGorderParametersPS}.
None of the tested models gives satisfactory agreement with
experiments for order parameters in headgroup $\alpha$ and $\beta$
carbons.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPS.pdf}
  \caption{\label{HGorderParametersPS}
    Order parameters for PS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are taken from experiments for POPS described in Supplementary Information.
    The vertical bars shown for the computational values are not error bars, but demonstrate that for these systems we had at least two data sets; the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average.
  }
  \todo{Check and report all the counterions.}
\end{figure}

@@

Glycerol backbone order parameters seems similar in all models,
except in Slipids. Even thought glycerol backbone order parameter
values are not yet experimentally available for PS lipids,
the comparison with the results for PC lipids suggest that
Slipid model do not correctly capture the glycerol backbone
structure \cite{botan15}. The glycerol backbone structures between PC and
PS lipids simulated with CHARMM36 are compared with the structures
simulated with CHARMM36 in Fig. \ref{glycerol_buslaev}.
The differences in sampled conformation leading to the order parameter
differences are clearly visible in the figure.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/glycerol_buslaev.png}
  \caption{\label{glycerol_buslaev}
    Snapshots overlayed from different simulations for glycerol backbone region
    by Pavel Buslaev.
  }
\end{figure}

@@

\subsection{Effect of PS on PC headgroup}
The headgroup order parameters of POPC 
mixed with PS lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
from different simulation model and experiments \cite{scherer87} with different
mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
This is not the case in simulation data shown in Fig. \ref{HGorderparametersPCvsPEPSPG}.
%The addition of DOPE into a POPC and DOPC bilayers significantly decreases the PC headgroup
%order parameters in simulations with OPLS compatible version of the Berger force field \cite{tieleman06}
%in contrast to experiments \cite{scherer87}. On the other hand, the increase of the PC
%headgroup order parameters in CHARMM36 simulations mixed with PS and PG lipids is significantly
%smaller than in experiments.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPS.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    PC headgroup order parameters from mixtures with PE, PS and PG
    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure}

@@

\subsection{Effect of PC on PS headgroup}
The headgroup order parameters of PS mixed with varying amounts of PC
from simulations and experiments \cite{borle85,roux90}
are shown in Fig. \ref{HGorderparametersPSPGvsPC}. The effect of increasing
amount of PC to PS headgroup seems to qualitatively incorrect in CHARMM36 simulations.
The $\beta$-carbon order parameter increases in experiment, but decreases
in simulations with both tested counterions (Na+ and K+). Larger
$\alpha$-carbon order parameter decreases with the addition of PC
in experiment, while the lower remains unchanged. In simulations the
larger increases and the lower decreases.
Interestingly, the $\alpha$-carbon order parameters are closer to experiments
in pure PS system with K+ counterions than with Na+. 
%The changes in PG headgroup order parameters are minor in simulations, which
%is in line with the only available experiment for the $\beta$-carbon.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPSvsPC.eps}
  \caption{\label{HGorderparametersPSPGvsPC}
    PS order parameters from mixtures with PC
    lipids with various mole fractions from different simulation models and experiments \cite{borle85,roux90}.
    Signs for PS are measures as described in SI.
  }
  \todo{Some simulations contain potassium as counterions, while some sodium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
\end{figure}

>>>>

\subsection{Headgroup and glycerol backbone structure in simulations and experiments of PS lipid bilayers}
The headgroup order parameters of DOPS and POPS bilayers 
from different experiments and simulations are shown in
Figs. \ref{HGorderParametersPS}.
None of the tested models gives satisfactory agreement with
experiments for order parameters in headgroup $\alpha$ and $\beta$
carbons.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPS.pdf}
  \caption{\label{HGorderParametersPS}
    Order parameters for PS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are taken from experiments for POPS described in Supplementary Information.
    The vertical bars shown for the computational values are not error bars, but demonstrate that for these systems we had at least two data sets; the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average.
  }
  \todo{Check and report all the counterions.}
\end{figure}

@@

Glycerol backbone order parameters seems similar in all models,
except in Slipids. 
Slipid model do not correctly capture the glycerol backbone
structure, as already observed previously for PC lipids \cite{botan15}.
The glycerol backbone structures between PC and
PS lipids simulated with Slipids are compared with the structures
simulated with CHARMM36 in Fig. \ref{glycerol_buslaev}.
The differences in sampled conformation leading to the order parameter
differences are clearly visible in the figure.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/glycerol_buslaev.png}
  \caption{\label{glycerol_buslaev}
    Snapshots overlayed from different simulations for glycerol backbone region
    by Pavel Buslaev.
  }
\end{figure}

@@

\subsection{Headgroup structure in PS and PC mixtures}
The headgroup order parameters of POPC 
mixed with PS lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
from different simulation model and experiments \cite{scherer87} with different
mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
This is not the case in simulation data shown in Fig. \ref{HGorderparametersPCvsPEPSPG}.
%The addition of DOPE into a POPC and DOPC bilayers significantly decreases the PC headgroup
%order parameters in simulations with OPLS compatible version of the Berger force field \cite{tieleman06}
%in contrast to experiments \cite{scherer87}. On the other hand, the increase of the PC
%headgroup order parameters in CHARMM36 simulations mixed with PS and PG lipids is significantly
%smaller than in experiments.
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPS.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    Headgroup order parameters from PC:PS mixtures from 
    different simulation models and experiments.
    Left panel shows the PC headgroup order parameters (experimental results from Ref. \citenum{scherer87},
    signs are determined as discussed in \cite{botan15,ollila16}).
    Rigth panel shows PS headgroup order parameters (experimental result for pure POPS measured in this
    work at 298K, experimental result for mixture from Ref. \citenum{roux90} at 298K).
    Counterions in expriments are sodium, while potassium is used in simulations.
    Using sodium in simulations do not have a significant effect.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.} \\
  \todo{We need results also from other than CHARMM36 force field.}
\end{figure*}

@@

%\subsection{Effect of PC on PS headgroup}
The headgroup order parameters of PS mixed with varying amounts of PC
from simulations and experiments \cite{borle85,roux90}
are shown in Fig. \ref{HGorderparametersPSPGvsPC}. The effect of increasing
amount of PC to PS headgroup seems to qualitatively incorrect in CHARMM36 simulations.
The $\beta$-carbon order parameter increases in experiment, but decreases
in simulations with both tested counterions (Na+ and K+). Larger
$\alpha$-carbon order parameter decreases with the addition of PC
in experiment, while the lower remains unchanged. In simulations the
larger increases and the lower decreases.
Interestingly, the $\alpha$-carbon order parameters are closer to experiments
in pure PS system with K+ counterions than with Na+. 
%The changes in PG headgroup order parameters are minor in simulations, which
%is in line with the only available experiment for the $\beta$-carbon.
%\begin{figure}[]
%  \centering
%  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPSvsPC.eps}
%  \caption{\label{HGorderparametersPSPGvsPC}
%    PS order parameters from mixtures with PC
%    lipids with various mole fractions from different simulation models and experiments \cite{borle85,roux90}.
%    Signs for PS are measures as described in SI.
%  }
%  \todo{Some simulations contain potassium as counterions, while some sodium.
%    All experiments here contain some amount of sodium salt. The best
%    ion concentrations for comparison should be figured out.} \\
%\end{figure}
